Parsing Postfix log files

### 2007-11-20 doc/logparser.tex
More XXX items dealt with.

@@

It is important to compare and contrast newly developed programs,
algorithms and parsers against those already available, to accurately judge
what, if any, improvements are delivered by the newcomers.  There are not
that many previously developed Postfix log parsers, indeed it was quite
difficult to find ten parsers to review for this project, and the
functionality offered ranges from quite basic to much more mature,
depending on the needs of the creator.  The programs are not reviewed from
an independent viewpoint; the objective is to compare and contrast each
program with \parsername{}.  One important difference between \parsername{}
and the parsers reviewed is that only \parsername{} makes it possible to
parse new log lines without modifying the core parsing algorithm.  Although
this may seem like a trivial point, is it substantially more difficult to
understand a program's entire parsing algorithm, identify the correct
location to change, and make the appropriate changes without adversely
affecting parsing, particularly as there may be edge cases which are not
immediately obvious.\footnote{See
section~\refwithpage{yet-more-aborted-delivery-attempts} for a complication
which occurs only four times in 93 log files tested.}  Requiring changes to
the parsing algorithm also complicates upgrades, as the changes must be
preserved, and may clash with changes made by the developer.  \parsername{}
allows the user to add new rules to the database without changing the
parsing algorithm, unless the new log lines to be parsed require
functionality not already provided by the algorithm.  If the new log lines
do require new functionality, new actions can be added to the parser
without modifying existing actions or other parts of the algorithm; only in
the rare case that the new actions require support from other sections of
the code will more extensive changes be required.

XXX ADD THE PREVIOUS TWO SENTENCES TO THE PARSING RULES/ALGORITHM
SECTION\@.

@@

        XXX TEST THIS\@: THE NEXT PARAGRAPH IS CRAP\@.

        Due to the paucity of data gathered by this tool the author has not
        experimented with it, but based on the documentation it is obvious
        that it doesn't attempt to gather the breadth of data required for
        in-depth analysis, and presumably doesn't attempt to correlate log
        lines per-connection or per-mail.

>>>>

XXX WHERE SHOULD THE NEXT PARAGRAPH BE PLACED\@?

\label{why separate rules and algorithm}

Separating the rules and algorithm makes it possible to parse new log lines
without modifying the core parsing algorithm.  Although this may seem like
a trivial point, is it substantially more difficult to understand a
program's entire parsing algorithm, identify the correct location to
change, and make the appropriate changes without adversely affecting
parsing, particularly as there may be edge cases which are not immediately
obvious.\footnote{See
section~\refwithpage{yet-more-aborted-delivery-attempts} for a complication
which occurs only four times in 93 log files tested.}  Requiring changes to
the parsing algorithm also complicates upgrades, as the changes must be
preserved during the upgrade, and may clash with changes made by the
developer.  \parsername{} allows the user to add new rules to the database
without changing the parsing algorithm, unless the new log lines to be
parsed require functionality not already provided by the algorithm.  If the
new log lines do require new functionality, new actions can be added to the
parser without modifying existing actions or other parts of the algorithm;
only in the rare case that the new actions require support from other
sections of the code will more extensive changes be required.

@@

It is important to compare and contrast newly developed programs,
algorithms and parsers against those already available, to accurately judge
what, if any, improvements are delivered by the newcomers.  There are not
that many previously developed Postfix log parsers, indeed it was quite
difficult to find ten parsers to review for this project, and the
functionality offered ranges from quite basic to much more mature,
depending on the needs of the creator.  The programs are not reviewed from
an independent viewpoint; the objective is to compare and contrast each
program with \parsername{}.  One important difference between \parsername{}
and the parsers reviewed is that only \parsername{} makes it possible to
parse new log lines without modifying the core parsing algorithm; see
section~\refwithpage{why separate rules and algorithm} for a discussion of
why the separation of rules from the algorithm is beneficial.

@@

        Isoqlog produces a report listing the number of mails sent by each
        unique sender address, and separately the total number of bytes
        transferred; both reports are produced for daily, monthly and
        annual time spans, but only for the domains listed in its
        configuration file.  It appears to ignore all log lines except for
        those for the current day, though it does maintain a record of data
        previously extracted, which the newly extracted data is merged into
        (no information is provided on the format of the data store).  It
        doesn't utilise rejection log lines in any way, so is unsuitable
        for the purposes of this project.  It doesn't handle any of the
        complications discussed in this document, doesn't gather the
        breadth of data required for this project, and ignores the majority
        of log lines produced by Postfix.
